### 2020-08-13 Manuscript/Response.to.reviewers.tex

\documentclass[11pt]{article}

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}
\usepackage{xcolor}

\begin{document}
\begin{linenumbers}

\textit{\tiny	This manuscript was compiled on \today }	
\section*{}

\textbf{ 
 Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters}

\textbf {Response to reviewers}

\vspace*{1em}

We are thankful for the feedback, comments and suggestions Our manuscript RSPB-2019-2901  "Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters" has received. Find below our response to their suggestions, and how we implement them in our manuscript.

\vspace*{1em}
\textbf{\large{ Reviewer 1}}:

\begin{itemize}

\item{1.      The manuscript needs to be restructured so that it follows a more conventional Intro, Methods, Results, Discussion format. In its current format, the results are difficult to follow as some of the methods that are crucial in interpreting the data are not detailed until the end of the MS. Please revise so that Methods come before Results. }

\textbf{Response:}

We have now reorganized the manuscript sections following a standard structure. As a result, a few sentences in the results section that introduced the procedures follow to model and project the likelihood of taxon presence have now been merged  into the Methods section:

line 107: Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.}

\item{2.      This is certainly not the first eDNA metabarcoding paper to track shifts in community composition across environmental gradients, yet there are almost no references to other studies that have used this approach, nor how eDNA metabarcoding is being applied more broadly as a monitoring tool. Pleas incorporate this literature into the Introduction. }

\textbf{Response:}

We now include in the Introduction references to previous efforts to use eDNA metabarcoding for community composition changes assessment linked to environmental gradients, and biomonitoring of marine fauna in general

Line 78: {\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}

\item{3.      The manuscript suffers from a lack of transparent results pertaining to the sequencing and bioinformatic methods. How many samples failed to pass QC? How many reads per sample were the analyses based on? How many samples did you collect from each site? I think this manuscript will benefit from including a table in the main text that summarizes some of these data.}

\textbf{Response:}

It was the intent of the authors to have raw data and scripts available for the reviewers and the readership when published. We have now created an open github repository with all the code and the raw fastq files are in a google drive link.

\item{4.      The title needs to be more specific as this study only focusses on plankton communities. Please consider including Plankton in the title to give the readers a better sense of what the paper is actually about.}

\textbf{Response:}

The reviewer is right in pointing out that the dataset used here is limited to  planktonic taxa. We have reworded our title, which now reads {\color{red} Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Plankton}

\item{5.      The figures do a good job at displaying the study site and data, however they are very busy and figure legends/axes are very difficult to read in print. Please consider revising figures to make them more legible and digestible. For example, perhaps Figure 3 only needs to include taxa that show significant shifts? }

\textbf{Response:}

We agree with the reviewer and we have rethought and redesigned figure 3 to highlight the taxa that shows meaningful shifts - focusing on the most varying diatoms and dinoflagellates we reveal the magnitude and repercussion of the changes in suitability. Panel C has been moved to the supplementary material

\end{itemize}

\vspace*{1em}
\textbf{\large{ Reviewer 2}}:

\begin{itemize}
\item{Figures: the colors red and blue are used both to distinguish time periods (2019/2095) and locality (San Juan/Hood Canal) in different figures. Figure 2 legend doesn?t say which color is which (presumably they are the same as in Figure 1). Use consistent colors for categories across figures and use different colors for time period and locality, if possible.}

\textbf{Response:}
    We realize that the shared color scales can lead to confusion. This has now been fixed, and we employ now a different color scales for temporal and spatial variables.

\item{Figure 2 is presented as for Hood Canal in summer but differs from the image in the supplemental information for that same place and time period. The points and axes in the CAP plot are the same, but the plots in sections B and C differ. What?s the difference between Fig 2 and the (ostensibly same) figure in the supplement? If these differ, how do we interpret the others in the supplement? The legend doesn?t explain the colors in Fig 2C.}

\textbf{Response:}

Figure 2 should be exactly the same as the first part of the Supplemental figure - A thorough review of the code that generates the supplemental figures revealed a small difference in the input dataset. We thank the reviewer for noticing this, and now the code that generates each of the figures is consistent.

\item{As Figure 3 is divided into sections A, B, and C, these ought to be referred to in the text (see below). The text is very small and hard to read in the review version I have.}

\textbf{Response:}

In an effort to declutter Figure 3, and following suggestions from reviewer 1, we have focused the figure in those species showing a more dramatic change in suitability. The new version of the figure has two panels (A and B) and we refer to them specifically in the text

\item{Lines 74-76 ?with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans?: This seems like an important point that perhaps should be included in the Introduction.}

\textbf{Response:}

Agreed. The advantage of the Hood Canal in this regard is paramount, and we now have included this concept in the introduction.  Line 52 onwards: {\color{red} [in close geographic proximity]. This gradients are pronounced and show values of pH and temperature expected worldwide by the end of the century}.

\item{Line 96: Specify Figure 2A}

\textbf{Response:}

TODO

\item{Line 99: The value of predicting which cluster (?community?) a given set of environmental parameters will fall within is not entirely clear to me. Is there a consistent set of taxa that make of up these clusters across time and space? Can specific groups of taxa be deduced from the ?community? number? Do the environmental variables accurately predict which set of taxa will be present? This section needs a bit more clarity.}

\textbf{Response:}

We think the relationship between clusters and the environmental variables are important in a few aspects. First, as a means of ground truthing our findings, it has great value as the clusters were detected blind to the environmental data, and the CAP analysis return the distinctiveness of these clusters in that space - Figure 2B highlights the translation of those modified axis of the CAP plot into the real values of temperature and pH. Second, these clusters (maybe communities is a more committed term that requires a longer dataset in which to confirm co-ocurrence of members) are a combination of species that are unique to them plus changes in relative presence - not a clear cut between total abence of a species but definitely a gradient in the relative importance of the species present.
We have reworded the section

TODO

\item{Line 102-104: According to my interpretation of Fig. 2C, it seems like the colder water communities actually had *lower* values of the given taxa, as the differences are all negative relative to other communities. It may help to specify in the text which characteristics are specific to which communities (communities 2 and 3 both span the lower temperature area of Fig. 2B).}

\item{Line 112-117: The authors base much of their calculation on a set of simulated temperature and pH values for different regions in the Salish Sea in 2095. The underlying data was not available so it?s not clear to me exactly how they subdivide the regional variation in the model predictions from Khangaonkar et al. (2019) (see comments on the Methods section as well).}

\item{Line 118-120: Please use figure subsections (A-C) when referring to Figure 3. Line 188 should reference Fig. 3A, line 119 should refernce Fig. 3C, and line 120 should reference Fig. 3B (the text erroneously refers to middle and bottom panels, when it ought to be bottom and middle panels, respectively.}

\item{Line 121: Refer to the relevant section of Figure 3 (A, B, or C)}

\item{Line 121-133: When referring to specific taxa as winners or losers, I was expecting to see them in Figure 3B, but not all of them were there. Or if it did appear, it was only at one site, but it was mentioned as being an overall winner or loser.}

\item{Line 122-123: According to Figure 3B, Thalassionema, Navicula, and Nitzschia appear to refer specifically to San Juan. If so, please state this.}

\item{Line 123: Coscinodiscus and Ditylum seem to refer to Hood Canal. This ought to specified as before.}

\item{Line 129: Write out Emiliania. The ?sp.? after Alexandrium should not be italicized. This sentence might read better if written: ?The coccolithophore Emilinania huxleyi and the dinoflagellate Alexandrium sp. both find more suitable habitat? (rather than ?each finds?)}

\item{Line 136-137: This is minor, but this section (or methods) might benefit from a brief mention of why ANOVA was used in one case and Wilcoxon in another.}

\item{Line 138-141: It seems like this paragraph could be moved to the discussion section.}

\item{Line 147-148: Need a citation for Coscinodiscus as key fish food.}

\item{Line 178: Spell out harmful algal bloom the first time it?s used.}

\item{Line 192: Might be helpful to specifically mention carbon sequestration as an important aspect of sinking rates.}

\item{Line 204-205: Citation needed for fish mortality. Eutrophication seems like a cause of blooms, not a result of them. What are the potential cascade effects on benthic-pelagic coupling? The term here on its own isn?t really saying anything.}

\item{Line 208-214: This is a big aspect of why this study is valuable! I think it?s worth discussing this in the introduction.}

\item{Line 237: Should be Scripps Institution of Oceanography}

\item{Line 241: Should be Figure 1C}

\item{Line 267-268: Which index of abundance from Kelly et al. (2019) was used? That paper presents several and though it declares one to be more useful, I think it would be helpful in this manuscript to state which one was used.}

\item{Line 273-282: It?s hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area. I think this section could use a bit of clarification, and all the code and raw data should be given (the provided github link is broken).}

\end{itemize}
 \end{linenumbers}
\end{document}

### 2020-08-13 Manuscript/What.lies.ahead.reviewed.tex
Added the manuscript folder

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}

@@

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions in each of the areas and seasons sampled, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliana huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.

@@

\section{Methods}

\subsection*{Sampling}
We collected water samples to assess eDNA communities in two regions of the Salish Sea (Washington, USA): San Juan Island and the Hood Canal. These sites experience substantial variation in water chemistry and other environmental conditions despite geographic proximity (ca. 300km; Fig. 1). We sampled eight sites monthly for approximately 1.5 years (March 2017 to August 2018), taking three 1L samples (biological replicates; ca. 10m apart) each month at each site (261 bottle samples total). Each sample was filtered through a 0.45 $\mu$m  cellulose filter, and the filter preserved in Longmire buffer until DNA extraction \cite{renshaw2015room}. Concurrently, we collected one 120 ml water sample from each site and poisoned it with 0.1 ml of saturated HgCl$_2$ for carbonate chemistry analysis, following \cite{riebesell2011guide}. We also collected \textit{in situ} measurements of temperature, salinity and dissolved oxygen using a handheld multiprobe (Hanna Instruments, USA) and a portable refractometer. We note that many unmeasured variables influence planktonic communities (e.g., nutrients, sunlight, wave energy), but that our set of measured parameters clearly distinguished communities and was adequate for our purposes.

@@

Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects), allowing slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. See Supporting Information for analytical code.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2019-11-06.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

>>>>

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}
\usepackage{xcolor}

@@

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions in both areas sampled, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

%% Emiliania huxleyi 
%%%%%%%
%%%%%%%

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.  
% Adding stuff on eDNA metabarcoding for community change.

{\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}

@@

%

\section{Methods}
\subsection*{Sampling}

We collected water samples to assess eDNA communities in two regions of the Salish Sea (Washington, USA): San Juan Island and the Hood Canal. These sites experience substantial variation in water chemistry and other environmental conditions despite geographic proximity (ca. 300km; Fig. 1). We sampled eight sites monthly for approximately 1.5 years (March 2017 to August 2018), taking three 1L samples (biological replicates; ca. 10m apart) each month at each site (261 bottle samples total). Each sample was filtered through a 0.45 $\mu$m  cellulose filter, and the filter preserved in Longmire buffer until DNA extraction \cite{renshaw2015room}. Concurrently, we collected one 120 ml water sample from each site and poisoned it with 0.1 ml of saturated HgCl$_2$ for carbonate chemistry analysis, following \cite{riebesell2011guide}. We also collected \textit{in situ} measurements of temperature, salinity and dissolved oxygen using a handheld multiprobe (Hanna Instruments, USA) and a portable refractometer. We note that many unmeasured variables influence planktonic communities (e.g., nutrients, sunlight, wave energy), but that our set of measured parameters clearly distinguished communities and was adequate for our purposes.

@@

Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. We allowed} slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. See Supporting Information for analytical code.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2019-11-06.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).
